---
title: ''
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(janitor)
library(sf)
library(tigris)
# library(scales)
library(extrafont)
library(patchwork)
# library(ggtext)
library(cowplot)
# extrafont::loadfonts()


# setwd("21-0119_Kenya_Census")
```

## Get Data

```{r}
tt <- tidytuesdayR::tt_load("2021", week = 4)

households <- 
  clean_names(tt$households) %>% 
  mutate(across(where(is_character), str_trim))
crops <- clean_names(tt$crops)
gender <- clean_names(tt$gender)

```

## Investigate

I'm curious in trying some mapping stuff, as I'm not super familiar/comfortable with it in R.  
Maybe create some sort of heat map by prevalence of households for each county in Kenya?

Do maps exist to this detail already in ggplot2/maps

```{r}
kenya_map <- map_data("world", region = "kenya") %>% as_tibble()


ggplot(kenya_map, aes(long, lat, group = group)) +
  geom_polygon(fill= "white", color = "black") + 
  coord_quickmap()


```

ggplot doesn't go to the level of counties, at least not that I could find.

Found a shapefile on arcgis.com (	http://www.arcgis.com/home/item.html?id=5f83ca29e5b849b8b05bc0b281ae27bc) via africaopendata.org(https://africaopendata.org/dataset/kenya-counties-shapefile)

```{r load shapefile}
kenya_sf <- 
  sf::read_sf("kenya_counties_shapefile/County.shp") %>% 
  mutate(Name = str_to_upper(Name))


# edit df so it joins correctly
kenya_sf <- 
  kenya_sf %>% 
  mutate(Name = if_else(str_detect(Name, "MURANG"), "MURANG'A", Name))
```

## Reduce Data

Let's create four heat maps, one for each of crop.

```{r crops reduce}
# fortunately, there's a "KENYA" total row

crops_names <- 
  crops %>% 
  filter(str_detect(sub_county, regex("kenya", ignore_case = TRUE))) %>% 
  select(-sub_county) %>% 
  pivot_longer(everything()) %>% 
  arrange(desc(value)) %>% 
  # remove "farming" as it's unclear what exactly that means.  Definition is "Population growing farming crops"
  filter(!str_detect(name, regex("farming", ignore_case = TRUE))) %>% 
  # head(5) %>% 
  pull(name) %>% 
  sort()

crops_names_format <- 
  crops_names %>% 
  str_replace("_", " ") %>% 
  str_to_title()

crop_values <- 
  crops %>% 
  # remove kenya total
  filter(!str_detect(sub_county, regex("kenya", ignore_case = TRUE))) %>% 
  select(sub_county, any_of(crops_names)) %>% 
  # mutate(across(.fns = ~ replace_na(., 0))) %>% 
  pivot_longer(!sub_county, 
               names_to = "crop") %>% 
  mutate(value = if_else(!is.na(value), value / 1000, value)) %>% 
  split(x = ., f = .$crop)


```


## Join and plot

```{r test}
# test with avocado

foo <- crop_values$avocado


foo <- geo_join(
  spatial_data = kenya_sf, 
  data_frame = foo, 
  by_sp = "Name", 
  by_df = "sub_county", 
  how = "inner"
)
```

```{r ggplot function}

kenya_ggplot <- 
  function(.data) {
    ggplot(data = .data) +
      geom_sf(aes(fill = value), 
              color = "white", 
              size = 0.2) + 
      coord_sf() + 
      scale_fill_viridis_c(option = "E", na.value = "light gray") + 
      labs(fill = "# HH (k)") + 
      theme_void() +
      theme(
        legend.text = element_text(family = "Consolas"), 
        plot.title = element_text(family = "Verdana", 
                                  size = 12), 
        
        # plot.margin = margin(20, 20, 20 , 20), 
        legend.position = "right") 
  }
```

## Map to create plots and save them

```{r full plots}

map2(.x = crop_values, 
     .y = crops_names_format,
     .f = ~ geo_join(spatial_data = kenya_sf, 
                     data_frame = .x, 
                     by_sp = "Name", 
                     by_df = "sub_county", 
                     how = "inner") %>% 
       {kenya_ggplot(.data = .) + 
           labs(title = as.character(.y))} %>% 
       assign(value = ., 
              x = str_c("kenya_plot_", str_to_lower(str_replace(as.character(.y), " ", "_"))), 
              envir = globalenv())
     
)
```


## combine plots to a grid

```{r}
grid <- 
  (kenya_plot_avocado | kenya_plot_cashew_nut | kenya_plot_citrus) / 
  (kenya_plot_coconut | kenya_plot_coffee | kenya_plot_khat_miraa) / 
  (kenya_plot_macadamia | kenya_plot_mango | kenya_plot_tea)

final_plot <- 
  grid +
  plot_layout(nrow = 3) +
  plot_annotation(
    title = "Permant Crop Production in Kenya", 
    subtitle = "\nBy county and Household Count\nData from the 2019 Kenya Population and Housing Census\n", 
    caption = "Plot by @greufek\nData from rKenyaCensus via Tidy Tuesday\nKenya Counties shapefile via openAfrica -- africaopendata.org/dataset/kenya-counties-shapefile") +  
  theme(
    # plot.margin = margin(10, 0, 25, 0), 
    plot.title = element_text(family = "Verdana", size = 16, hjust = 0),
    plot.subtitle = element_text(family = "Verdana", size = 12, hjust =  0), 
    plot.caption = element_text(family = "Verdana", hjust = 0.95, size = 8) 
  )
```

## Save plot 

```{r save plot}
dte <- Sys.Date()

ggsave(plot = final_plot,
       filename = str_c("output/", dte, "_kenya_census_map_crops.jpeg"),
       dpi = "retina",
       device = "jpeg", 
       width = 7, 
       height = 9, 
       units = "in")
```


